Extending statistical models for source attribution of zoonotic diseases: A study of campylobacteriosis

### 2018-03-27 paper/finalish/article.tex
utilised -> utilized

\item Compute the acceptance probability $\alpha$, a product of the likelihood ratio, the proposal ratio and the prior ratio, given by
\begin{align*}
\alpha=\frac{L\Big(F^*; \text{ST}\Big)}{L\Big(F; \text{ST}\Big)}\frac{Q\Big(\theta_{(t)} \vert \theta_{(t)}^*\Big)}{Q\Big(\theta_{(t)}^* \vert \theta_{(t)}\Big)}\frac{p\Big(\theta_{(t)}^*\Big)}{p\Big(\theta_{(t)}\Big)},
\end{align*}
where the likelihood $L\Big(F^*; \text{ST}\Big)$ is given by equation \eqref{Olikh}. \item Sample $u$ from $U(0,1)$. If $u < \min \left\{1, \alpha \right\}$, the proposals $f^{(*)}$, $F^{(*)}$ and $L\Big(F^{(*)}; \text{ST}\Big)$ are utilised for the next iteration; otherwise, the previous values are retained.
\end{enumerate}
\item Repeat from step \ref{algorithm_loop} for the given number of iterations.

@@

As noted previously, a major public health initiative in 2007 led to a significant reduction in the number of cases of campylobacteriosis in New Zealand. In order to examine the effects of this change on attribution probabilities, we repeated the analysis by including an interaction with time period (2005--2007 and 2008--2014). Figure~\ref{fig:intervention} shows that the general trend in attribution by rurality for each of the time periods utilising a linear trend on the logit scale to incorporate rurality. There is a clear difference, with a significantly lower attribution to poultry (and correspondingly higher attribution to ruminants) in all but the most rural of areas, being strongest in highly urban areas. Thus, although the intervention did not eliminate infection arising from poultry \cite{MuelM}, the reduction highlights the significant improvement in contribution of poultry to the disease, particularly in urban areas where most cases lived.

\section{Discussion}
Models that determine the source of human infection, particularly for zoonotic pathogens that originate in animal populations, are of considerable value to public health policy makers. However, such models may be complex, particularly when utilising evolutionary models. An outstanding question is whether such complexity is required, or whether a simpler model may work as effectively. Here we developed a relatively simple model to estimate the attribution probability for each source of \textit{Campylobacter} infection. This model differs to the asymmetric Island model, in that it does not model pathogenic evolution, opting instead to use infer the sampling distribution of genotypes directly from the observed count data.

>>>>

\item Compute the acceptance probability $\alpha$, a product of the likelihood ratio, the proposal ratio and the prior ratio, given by
\begin{align*}
\alpha=\frac{L\Big(F^*; \text{ST}\Big)}{L\Big(F; \text{ST}\Big)}\frac{Q\Big(\theta_{(t)} \vert \theta_{(t)}^*\Big)}{Q\Big(\theta_{(t)}^* \vert \theta_{(t)}\Big)}\frac{p\Big(\theta_{(t)}^*\Big)}{p\Big(\theta_{(t)}\Big)},
\end{align*}
where the likelihood $L\Big(F^*; \text{ST}\Big)$ is given by equation \eqref{Olikh}. \item Sample $u$ from $U(0,1)$. If $u < \min \left\{1, \alpha \right\}$, the proposals $f^{(*)}$, $F^{(*)}$ and $L\Big(F^{(*)}; \text{ST}\Big)$ are utilized for the next iteration; otherwise, the previous values are retained.
\end{enumerate}
\item Repeat from step \ref{algorithm_loop} for the given number of iterations.

@@

As noted previously, a major public health initiative in 2007 led to a significant reduction in the number of cases of campylobacteriosis in New Zealand. In order to examine the effects of this change on attribution probabilities, we repeated the analysis by including an interaction with time period (2005--2007 and 2008--2014). Figure~\ref{fig:intervention} shows that the general trend in attribution by rurality for each of the time periods utilizing a linear trend on the logit scale to incorporate rurality. There is a clear difference, with a significantly lower attribution to poultry (and correspondingly higher attribution to ruminants) in all but the most rural of areas, being strongest in highly urban areas. Thus, although the intervention did not eliminate infection arising from poultry \cite{MuelM}, the reduction highlights the significant improvement in contribution of poultry to the disease, particularly in urban areas where most cases lived.

\section{Discussion}
Models that determine the source of human infection, particularly for zoonotic pathogens that originate in animal populations, are of considerable value to public health policy makers. However, such models may be complex, particularly when utilizing evolutionary models. An outstanding question is whether such complexity is required, or whether a simpler model may work as effectively. Here we developed a relatively simple model to estimate the attribution probability for each source of \textit{Campylobacter} infection. This model differs to the asymmetric Island model, in that it does not model pathogenic evolution, opting instead to use infer the sampling distribution of genotypes directly from the observed count data.
